A Parallel, Backjumping Subgraph Isomorphism Algorithm using Supplemental Graphs

### 2015-08-26 talk/talk.tex
Colours

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}

\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.5cm, font=\small\bfseries}}
\tikzset{notvertex/.style={vertex, color=white, text=black}}
\tikzset{plainvertex/.style={vertex}}
\tikzset{vertexc1/.style={vertex, fill=uofgthistle}}
\tikzset{vertexc2/.style={vertex, fill=uofgpumpkin}}
\tikzset{vertexc3/.style={vertex, fill=uofgpillarbox}}
\tikzset{vertexc4/.style={vertex, fill=uofgturquoise}}
\tikzset{edge/.style={color=black!50!white}}
\tikzset{bedge/.style={ultra thick}}
\tikzset{edged/.style={color=screengrey, dashed}}
\tikzset{edgel3/.style={color=uofgturquoise, ultra thick}}

@@

\begin{frame}{Selected Previous Work}
    \begin{itemize}
        \item \textcolor{uofgthistle}{VF2}: backtracking search, interesting heuristics. Widely used. \\[0.1cm]
            {\scriptsize A (Sub)Graph Isomorphism Algorithm for Matching Large Graphs. \emph{Luigi P.
                Cordella, Pasquale Foggia, Carlo Sansone and Mario Vento}. {IEEE} Trans.
        Pattern Anal. Mach. Intell., 2004.}

        \item \textcolor{uofgthistle}{LAD}: locally all-different, and neighbourhood degree sequences. \\[0.1cm]
            {\scriptsize AllDifferent-based filtering for subgraph isomorphism. \emph{Christine
                Solnon}.  Artif. Intell., 2010.}

        \item \textcolor{uofgthistle}{SND}: distance-based filtering. \\[0.1cm]
            {\scriptsize Scoring-Based Neighborhood Dominance for the Subgraph Isomorphism Problem.
                \emph{Gilles Audemard, Christophe Lecoutre, Mouny Samy Modeliar, Gilles
            Goncalves and Daniel Porumbel}. CP 2014.}
    \end{itemize}
\end{frame}

@@

        \item When LAD and SND fail, they often manage less than \textcolor{uofgthistle}{one
            recursive call per second}, particularly with larger target graphs.
    \end{itemize}
\end{frame}

@@

\begin{frame}{Our Approach}
    \begin{itemize}
        \item \textcolor{uofgthistle}{Cheaper inference}: $10^4$ to $10^6$ recursive calls per second.
        \item Expensive \textcolor{uofgthistle}{preprocessing once} at the top of search, rather
            than computing distances and degree sequences during search.
        \item FC-CBJ instead of MAC:
            \begin{itemize}
                \item A \textcolor{uofgthistle}{counting-based all-different} propagator, which does more
                    than pairwise-$\ne$ AC but less than GAC.
                \item CBJ \textcolor{uofgthistle}{without explicit conflict sets}, and without modifying propagators.
            \end{itemize}
        \item \textcolor{uofgthistle}{Bit-parallelism} for all major operations.
        \item \textcolor{uofgthistle}{Thread-parallel} search:
            \begin{itemize}
                \item Safe, deterministic search via nested parallel folds.
                \item Explicit early work-stealing, to fake discrepancy search.
            \end{itemize}
    \end{itemize}
\end{frame}

@@

        \centering
        \begin{tikzpicture}
            \node [vertex] (Pc) at (0, 0) { c };
            \node [vertexc2] (Pd) at (1, 0) { d };
            \node [vertex] (Pb) at (1, 1) { b };
            \node [vertexc1] (Pa) at (0, 1) { a };

@@

            \node [vertexc1] (T1) at (4.13, 0.5) { 1 };
            \node [vertex] (T5) at (5, 0) { 5 };
            \node [vertex] (T6) at (6, 0) { 6 };
            \node [vertex] (T3) at (6, 1) { 3 };
            \node [vertex] (T2) at (5, 1) { 2 };
            \node [vertexc2] (T4) at (6.87, 0.5) { 4 };

@@

    \only<3> {
        \begin{itemize}
            \item We can do something stronger: rather than looking at distances, we can look at
                \textcolor{uofgthistle}{(simple) paths}, and we can count how many there are.

            \item This is \NP-hard in general, but only \textcolor{uofgthistle}{lengths 2 and 3} and
                counts of 2 and 3 are useful in practice.

            \item We construct these graph pairs \textcolor{uofgthistle}{once, at the top of
                search}, and use them for degree-based filtering at the top of search, and
                ``adjacency'' filtering during search.
        \end{itemize}
    }
\end{frame}

\begin{frame}{Counting-Based All-Different}
    \begin{itemize}
        \item \textcolor{uofgthistle}{One pass} through the variables, from smallest domain to largest.

        \item Track \textcolor{uofgthistle}{unions of domains} as we go along.

        \item Eliminate \textcolor{uofgthistle}{any detected Hall sets} from future variables.

        \item This \textcolor{uofgthistle}{can miss some deletions}, but has a very fast bitset
            implementation for large domains.
    \end{itemize}
\end{frame}

\begin{frame}{Backjumping}
    \begin{itemize}
        \item Rather than backtracking on failure, we can sometimes show it's safe to jump back
            several levels immediately.
        \item This can be done with without explicit conflict sets, and without modifying
            propagators.
        \item Producing smaller conflict sets from a failed all-different gives better results.
    \end{itemize}
\end{frame}

@@

            \tikzstyle{t} = [draw, fill, fill=uofgcobalt, rounded corners];
            \foreach \na in {1, ..., 3}{
                \foreach \nb in {1, ..., 3}{
                    \draw [t] (N\na\nb) -- (N\na\nb t1) -- (N\na\nb t2) -- cycle;
                }
            }

            \tikzstyle{c} = [draw, circle, fill, fill=uofgcobalt];
            \node [c] at (N) { };

@@

    \only<2> {
        \begin{itemize}
            \item Explicit early-first work stealing, to offset incorrect decisions early in search.
                \begin{itemize}
                    \item There are EHPs, and this gets rid of them more cheaply than discrepancy
                        searches.
                \end{itemize}
            \item Backjumping is a nested parallel fold with left-zero elements.
            \item Speculative, so we should not expect linear speedups (even on unsat instances).
        \end{itemize}
    }
\end{frame}

\begin{frame}{Sidenote on Safety, Parallelism, and Benchmarking}
    \begin{itemize}
        \item The way we do parallelism guarantees:
            \begin{itemize}
                \item Parallel search will never be substantially worse than sequential search.
                \item Adding more cores will never make things substantially worse (excluding
                    hardware weirdness).
                \item Running it twice will give more or less the same runtimes.
            \end{itemize}
        \item All comparisons are against a dedicated sequential implementation of a strong
            algorithm, not a threaded implementation run with one thread.
    \end{itemize}
\end{frame}

>>>>

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.2, 0.129412}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.5cm, font=\small\bfseries}}
\tikzset{notvertex/.style={vertex, color=white, text=black}}
\tikzset{plainvertex/.style={vertex}}
\tikzset{vertexc1/.style={vertex, fill=uofgburgundy, text=white}}
\tikzset{vertexc2/.style={vertex, fill=uofgsandstone, text=white}}
\tikzset{vertexc3/.style={vertex, fill=uofgforest, text=white}}
\tikzset{vertexc4/.style={vertex, fill=uofgheather, text=white}}
\tikzset{edge/.style={color=black!50!white}}
\tikzset{bedge/.style={ultra thick}}
\tikzset{edged/.style={color=screengrey, dashed}}
\tikzset{edgel3/.style={color=uofgrose, ultra thick}}

@@

\begin{frame}{Selected Previous Work}
    \begin{itemize}
        \item \textcolor{uofgrose}{VF2}: backtracking search, interesting heuristics. Widely used. \\[0.1cm]
            {\scriptsize A (Sub)Graph Isomorphism Algorithm for Matching Large Graphs. \emph{Luigi P.
                Cordella, Pasquale Foggia, Carlo Sansone and Mario Vento}. {IEEE} Trans.
        Pattern Anal. Mach. Intell., 2004.}

        \item \textcolor{uofgrose}{LAD}: locally all-different, and neighbourhood degree sequences. \\[0.1cm]
            {\scriptsize AllDifferent-based filtering for subgraph isomorphism. \emph{Christine
                Solnon}.  Artif. Intell., 2010.}

        \item \textcolor{uofgrose}{SND}: distance-based filtering. \\[0.1cm]
            {\scriptsize Scoring-Based Neighborhood Dominance for the Subgraph Isomorphism Problem.
                \emph{Gilles Audemard, Christophe Lecoutre, Mouny Samy Modeliar, Gilles
            Goncalves and Daniel Porumbel}. CP 2014.}
    \end{itemize}
\end{frame}

@@

        \item When LAD and SND fail, they often manage less than \textcolor{uofgrose}{one
            recursive call per second}, particularly with larger target graphs.
    \end{itemize}
\end{frame}

@@

\begin{frame}{Our Approach}
    \begin{itemize}
        \item \textcolor{uofgrose}{Cheaper inference}: $10^4$ to $10^6$ recursive calls per second.
        \item Expensive \textcolor{uofgrose}{preprocessing once} at the top of search, rather
            than computing distances and degree sequences during search.
        \item FC-CBJ instead of MAC:
            \begin{itemize}
                \item A \textcolor{uofgrose}{counting-based all-different} propagator, which does more
                    than pairwise-$\ne$ AC but less than GAC.
                \item CBJ \textcolor{uofgrose}{without explicit conflict sets}, and without modifying propagators.
            \end{itemize}
        \item \textcolor{uofgrose}{Bit-parallelism} for all major operations.
        \item \textcolor{uofgrose}{Thread-parallel} search:
            \begin{itemize}
                \item Safe, deterministic search via nested parallel folds.
                \item Explicit early work-stealing, to fake discrepancy search.
            \end{itemize}
    \end{itemize}
\end{frame}

@@

        \centering
        \begin{tikzpicture}
            \node [vertex] (Pc) at (0, 0) { c };
            \node [vertexc3] (Pd) at (1, 0) { d };
            \node [vertex] (Pb) at (1, 1) { b };
            \node [vertexc1] (Pa) at (0, 1) { a };

@@

            \node [vertexc1] (T1) at (4.13, 0.5) { 1 };
            \node [vertex] (T5) at (5, 0) { 5 };
            \node [vertex] (T6) at (6, 0) { 6 };
            \node [vertex] (T3) at (6, 1) { 3 };
            \node [vertex] (T2) at (5, 1) { 2 };
            \node [vertexc3] (T4) at (6.87, 0.5) { 4 };

@@

    \only<3> {
        \begin{itemize}
            \item We can do something stronger: rather than looking at distances, we can look at
                \textcolor{uofgrose}{(simple) paths}, and we can count how many there are.

            \item This is \NP-hard in general, but only \textcolor{uofgrose}{lengths 2 and 3} and
                counts of 2 and 3 are useful in practice.

            \item We construct these graph pairs \textcolor{uofgrose}{once, at the top of
                search}, and use them for degree-based filtering at the top of search, and
                ``adjacency'' filtering during search.
        \end{itemize}
    }
\end{frame}

\begin{frame}{Counting-Based All-Different}
    \begin{itemize}
        \item \textcolor{uofgrose}{One pass} through the variables, from smallest domain to largest.

        \item Track \textcolor{uofgrose}{unions of domains} as we go along.

        \item Eliminate \textcolor{uofgrose}{any detected Hall sets} from future variables.

        \item This \textcolor{uofgrose}{can miss some deletions}, but has a very fast bitset
            implementation for large domains.
    \end{itemize}
\end{frame}

\begin{frame}{Backjumping}
    \begin{itemize}
        \item Rather than backtracking on failure, we can sometimes show it's safe to
            \textcolor{uofgrose}{jump back several levels} immediately.
        \item This can be done with without explicit conflict sets, and without modifying
            propagators.
        \item Producing smaller conflict sets from a failed all-different gives better results.
    \end{itemize}
\end{frame}

@@

            \tikzstyle{t} = [draw, fill, fill=uofgrose, rounded corners];
            \foreach \na in {1, ..., 3}{
                \foreach \nb in {1, ..., 3}{
                    \draw [t] (N\na\nb) -- (N\na\nb t1) -- (N\na\nb t2) -- cycle;
                }
            }

            \tikzstyle{c} = [draw, circle, fill, fill=uofgrose];
            \node [c] at (N) { };

@@

    \only<2> {
        \begin{itemize}
            \item Explicit \textcolor{uofgrose}{early-first work stealing}, to offset incorrect
                decisions early in search.
                \begin{itemize}
                    \item There are EHPs, and this gets rid of them more cheaply than discrepancy
                        searches.
                \end{itemize}
            \item Backjumping is a nested parallel fold with left-zero elements.
            \item Speculative, so we should \textcolor{uofgrose}{not expect linear speedups} (even
                on unsat instances).
        \end{itemize}
    }
\end{frame}

\begin{frame}{Sidenote on Safety, Parallelism, and Benchmarking}
    \begin{itemize}
        \item The way we do parallelism guarantees:
            \begin{itemize}
                \item Parallel search will \textcolor{uofgrose}{never be substantially worse} than
                    sequential search.
                \item Adding more cores will never make things substantially worse (excluding
                    hardware weirdness).
                \item Running it twice will give more or less the same runtimes.
            \end{itemize}
        \item All comparisons are against a dedicated sequential implementation of a strong
            algorithm, not a threaded implementation run with one thread.
    \end{itemize}
\end{frame}
